---
title: "KNN to predict the GRADE (no pass/pass)"
author: "Bogdan Tanasa"
date: "L2"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### **THE SECTIONS in the RMARKDOWN DOCUMENT :**

<p>&nbsp;</p>

#### **1. INTRODUCTION**

<p>&nbsp;</p>

#### **2. DATA EXPLORATION**

<p>&nbsp;</p>

#### **3. DATA FILTERING**

<p>&nbsp;</p>

#### **4. DATA TRANSFORMATION**

<p>&nbsp;</p>

#### **5. TRAINING AND TEST SETS**

<p>&nbsp;</p>

#### **6. PRE-PROCESSING THE DATA**

<p>&nbsp;</p>

#### **7. PERFORMING THE TRAINING**

<p>&nbsp;</p>

#### **8. MAKING THE PREDICTIONS**

<p>&nbsp;</p>

#### **9. THE CONFUSION MATRIX**

<p>&nbsp;</p>

<p>&nbsp;</p>
#### **1. INTRODUCTION**
<p>&nbsp;</p>

We are using the data from **UCI** : !( https://archive.ics.uci.edu/ml/datasets/Student+Performance )

We are reading a file about **STUDENTS**, and we aim to predict whether they have passed or not the exams **(PASS/no_PASS)**;

The attributes in the **INPUT FILE** are the following : 

* 1 school - student's school (binary: "GP" - Gabriel Pereira or "MS" - Mousinho da Silveira)

* 2 sex - student's sex (binary: "F" - female or "M" - male)

* 3 age - student's age (numeric: from 15 to 22)

* 4 address - student's home address type (binary: "U" - urban or "R" - rural)

* 5 famsize - family size (binary: "LE3" - less or equal to 3 or "GT3" - greater than 3)

* 6 Pstatus - parent's cohabitation status (binary: "T" - living together or "A" - apart)

* 7 Medu - mother's education (numeric: 0 - none,  1 - primary education (4th grade), 2 - 5th to 9th grade, 3 - secondary education or 4  - higher education)

* 8 Fedu - father's education (numeric: 0 - none,  1 - primary education (4th grade), 2 - 5th to 9th grade, 3 - secondary education or 4 - higher education)

* 9 Mjob - mother's job (nominal: "teacher", "health" care related, civil "services" (e.g. administrative or police), "at_home" or "other")

* 10 Fjob - father's job (nominal: "teacher", "health" care related, civil "services" (e.g. administrative or police), "at_home" or "other")

* 11 reason - reason to choose this school (nominal: close to "home", school "reputation", "course" preference or "other")

* 12 guardian - student's guardian (nominal: "mother", "father" or "other")

* 13 traveltime - home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)

* 14 studytime - weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)

* 15 failures - number of past class failures (numeric: n if 1<=n<3, else 4)

* 16 schoolsup - extra educational support (binary: yes or no)

* 17 famsup - family educational support (binary: yes or no)

* 18 paid - extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)

* 19 activities - extra-curricular activities (binary: yes or no)

* 20 nursery - attended nursery school (binary: yes or no)

* 21 higher - wants to take higher education (binary: yes or no)

* 22 internet - Internet access at home (binary: yes or no)

* 23 romantic - with a romantic relationship (binary: yes or no)

* 24 famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent)

* 25 freetime - free time after school (numeric: from 1 - very low to 5 - very high)

* 26 goout - going out with friends (numeric: from 1 - very low to 5 - very high)

* 27 Dalc - workday alcohol consumption (numeric: from 1 - very low to 5 - very high)

* 28 Walc - weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)

* 29 health - current health status (numeric: from 1 - very bad to 5 - very good)

* 30 absences - number of school absences (numeric: from 0 to 93)

<p>&nbsp;</p>
#### **2. DATA EXPLORATION**
<p>&nbsp;</p>

```{r }

library(ggplot2)
library(reshape2)
library(readxl)
library(dplyr)

library(tibble)

library(class)
library(gmodels)
library(caret)
library(e1071)

######################################################
######################################################
######################################################
######################################################

FILE1="student.mat.txt"
# FILE2="student.por.txt"
# FILE3="student.mat.and.por.txt"

######################################################
######################################################
######################################################
######################################################

student <- read.delim(FILE1, sep="\t", header=T, stringsAsFactors=F)

######################################################
######################################################

summary(student)
str(student)
class(student)
```

Here we are starting to display the data for visual exploration.

```{r }
############################################################################################################
############################################################################################################
# 1 school - student's school (binary: "GP" - Gabriel Pereira or "MS" - Mousinho da Silveira)

unique(student$school)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=school, fill=school))

ggsave("display.1.school.png")

student$school = as.factor(student$school)

############################################################################################################
############################################################################################################
# 2 sex - student's sex (binary: "F" - female or "M" - male)

unique(student$sex)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=sex , fill=sex))

ggsave("display.2.sex.png")

student$sex = as.factor(student$sex)

############################################################################################################
############################################################################################################
# 3 age - student's age (numeric: from 15 to 22)

unique(student$age)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=age , fill=age))

ggsave("display.3.age.png")

# AGE is already on the numerical scale !!  
student$age = as.integer(student$age)

############################################################################################################
############################################################################################################
# 4 address - student's home address type (binary: "U" - urban or "R" - rural)

unique(student$address) ## [1] "U" "R"

ggplot(data = student) + 
       geom_bar(mapping = aes(x=address, fill=address))

ggsave("display.4.address.png")

student$address = as.factor(student$address)

############################################################################################################
############################################################################################################
# 5 famsize - family size (binary: "LE3" - less or equal to 3 or "GT3" - greater than 3)

unique(student$famsize)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=famsize, fill=famsize))

ggsave("display.5.famsize.png")

student$famsize = as.factor(student$famsize)

############################################################################################################
############################################################################################################
# 6 Pstatus - parent's cohabitation status (binary: "T" - living together or "A" - apart)

unique(student$Pstatus)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Pstatus, fill=Pstatus))

ggsave("display.6.Pstatus.png")

student$Pstatus = as.factor(student$Pstatus)

############################################################################################################
############################################################################################################
# 7 Medu - mother's education (numeric: 0 - none,  1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)

unique(student$Medu)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Medu, fill=Medu))

ggsave("display.7.Medu.png")

# we may wanna use the numerical values in various regression models
student$Medu = as.integer(student$Medu)

############################################################################################################
############################################################################################################
# 8 Fedu - father's education (numeric: 0 - none,  1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)

unique(student$Fedu)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Fedu, fill=Fedu))

ggsave("display.8.Fedu.png")

# we may wanna use the numerical values in various regression models
student$Fedu = as.integer(student$Fedu)

############################################################################################################
############################################################################################################
# 9 Mjob - mother's job (nominal: "teacher", "health" care related, civil "services" (e.g. administrative or police), "at_home" or "other")

unique(student$Mjob)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Mjob, fill=Mjob))

ggsave("display.9.Mjob.png")

student$Mjob = as.factor(student$Mjob)

############################################################################################################
############################################################################################################
# 10 Fjob - father's job (nominal: "teacher", "health" care related, civil "services" (e.g. administrative or police), "at_home" or "other")

unique(student$Fjob)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Fjob, fill=Fjob))

ggsave("display.10.Fjob.png")

student$Fjob = as.factor(student$Fjob)

############################################################################################################
############################################################################################################
# 11 reason - reason to choose this school (nominal: close to "home", school "reputation", "course" preference or "other")

unique(student$reason)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=reason, fill=reason))

ggsave("display.11.reason.png")

student$reason = as.factor(student$reason)

############################################################################################################
############################################################################################################
# 12 guardian - student's guardian (nominal: "mother", "father" or "other")

unique(student$guardian)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=guardian, fill=guardian))

ggsave("display.12.guardian.png")

student$guardian = as.factor(student$guardian)

############################################################################################################
############################################################################################################
# 13 traveltime - home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)

unique(student$traveltime)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=traveltime, fill=traveltime))

ggsave("display.13.traveltime.png")

# we may wanna use the NUMERICAL VALUES :
student$traveltime = as.integer(student$traveltime)

############################################################################################################
############################################################################################################
# 14 studytime - weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)

unique(student$studytime)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=studytime, fill=studytime))

ggsave("display.14.studytime.png")

# we may wanna use the NUMERICAL VALUES :
student$studytime = as.integer(student$studytime)

############################################################################################################
############################################################################################################
# 15 failures - number of past class failures (numeric: n if 1<=n<3, else 4)

unique(student$failures)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=failures, fill=failures))

ggsave("display.15.failures.png")

# we may wanna use the NUMERICAL VALUES :
student$failures = as.integer(student$failures)

############################################################################################################
############################################################################################################
# 16 schoolsup - extra educational support (binary: yes or no)

unique(student$schoolsup)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=schoolsup, fill=schoolsup))

ggsave("display.16.schoolsup.png")

student$schoolsup = as.factor(student$schoolsup)

############################################################################################################
############################################################################################################
# 17 famsup - family educational support (binary: yes or no)

unique(student$famsup)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=famsup, fill=famsup))

ggsave("display.17.famsup.png")

student$famsup = as.factor(student$famsup)

############################################################################################################
############################################################################################################
# 18 paid - extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)

unique(student$paid)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=paid, fill=paid))

ggsave("display.18.paid.png")

student$paid = as.factor(student$paid)

############################################################################################################
############################################################################################################
# 19 activities - extra-curricular activities (binary: yes or no)

unique(student$activities)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=activities, fill=activities))

ggsave("display.19.activities.png")

student$activities = as.factor(student$activities)

############################################################################################################
############################################################################################################
# 20 nursery - attended nursery school (binary: yes or no)

unique(student$nursery)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=nursery, fill=nursery))

ggsave("display.20.nursery.png")

student$nursery = as.factor(student$nursery)

############################################################################################################
############################################################################################################
# 21 higher - wants to take higher education (binary: yes or no)

unique(student$higher)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=higher, fill=higher))

ggsave("display.21.higher.png")

student$higher = as.factor(student$higher)

############################################################################################################
############################################################################################################
# 22 internet - Internet access at home (binary: yes or no)

unique(student$internet)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=internet, fill=internet))

ggsave("display.22.internet.png")

student$internet = as.factor(student$internet)

############################################################################################################
############################################################################################################
# 23 romantic - with a romantic relationship (binary: yes or no)

unique(student$romantic)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=romantic, fill=romantic))

ggsave("display.23.romantic.png")

student$romantic = as.factor(student$romantic)

############################################################################################################
############################################################################################################
# 24 famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent)

unique(student$famrel)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=famrel, fill=famrel))

ggsave("display.24.famrel.png")

# i believe that we can keep these as numerical :
student$famrel = as.integer(student$famrel)

############################################################################################################
############################################################################################################
# 25 freetime - free time after school (numeric: from 1 - very low to 5 - very high)

unique(student$freetime)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=freetime, fill=freetime))

ggsave("display.25.freetime.png")

# i believe that we can keep these as numerical :
student$freetime = as.integer(student$freetime)

############################################################################################################
############################################################################################################
# 26 goout - going out with friends (numeric: from 1 - very low to 5 - very high)

unique(student$goout)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=goout, fill=goout))

ggsave("display.26.goout.png")

# i believe that we can keep these as numerical :
student$goout = as.integer(student$goout)

############################################################################################################
############################################################################################################
# 27 Dalc - workday alcohol consumption (numeric: from 1 - very low to 5 - very high)

unique(student$Dalc)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Dalc, fill=Dalc))

ggsave("display.27.Dalc.png")

# i believe that we can keep these as numerical :
student$Dalc = as.integer(student$Dalc)

############################################################################################################
############################################################################################################
# 28 Walc - weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)

unique(student$Walc)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=Walc, fill=Walc))

ggsave("display.28.Walc.png")

# i believe that we can keep these as numerical :
student$Walc = as.integer(student$Walc)

############################################################################################################
############################################################################################################
# 29 health - current health status (numeric: from 1 - very bad to 5 - very good)

unique(student$health)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=health, fill=health))

ggsave("display.29.health.png")

# i believe that we can keep these as numerical :
student$health = as.integer(student$health)

############################################################################################################
############################################################################################################
# 30 absences - number of school absences (numeric: from 0 to 93)

unique(student$absences)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=absences, fill=absences))

ggsave("display.30.absences.png")

# i believe that we can keep these as numerical :
student$absences = as.integer(student$absences)

############################################################################################################
############################################################################################################
# $ G1        : int  5 5 7 15 6 15 12 6 16 14 ...

unique(student$G1)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=G1, fill=G1))

ggsave("display.0.G1.png")

# i believe that we can keep these as numerical, although we may not need it :
student$G1 = as.factor(student$G1)

############################################################################################################
############################################################################################################
# $ G2        : int  6 5 8 14 10 15 12 5 18 15 ...

unique(student$G2)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=G2, fill=G2))

ggsave("display.0.G2.png")

# i believe that we can keep these as numerical, although we may not need it :
student$G2 = as.factor(student$G2)

############################################################################################################
############################################################################################################
# $ G3        : int  6 6 10 15 10 15 11 6 19 15 ...

unique(student$G3)

ggplot(data = student) + 
       geom_bar(mapping = aes(x=G3, fill=G3))

ggsave("display.0.G3.png")

# i believe that we can covert it into RANGES of VALUES :
student$G3 = as.factor(student$G3)

############################################################################################################
############################################################################################################
############################################################################################################
############################################################################################################
############################################################################################################
############################################################################################################

summary(student)
str(student)
class(student)
```

<p>&nbsp;</p>
#### **3. DATA FILTERING**
<p>&nbsp;</p>

```{r }
## the OUTPUT VARIABLES is G3
## we may remove G1 and G2
## and other features that are nit numerical

student1 <- subset(student, select = -c(G1, G2))

student2 <- subset(student1, 
                   select = -c(school, sex, address, famsize, Pstatus, 
                   Mjob, Fjob, reason, guardian, schoolsup, famsup, paid, activities, nursery, 
                   higher, internet, romantic))

str(student2)

student2$G3 = as.factor(student2$G3)

table(student2$G3)

### for simplicity, to work with a copy of STUDENT3, let's call it STUDENT3

student3 = subset(student2, 
                  select= c(age, traveltime, studytime, failures, absences, G3))

table(student3$G3)
```

<p>&nbsp;</p>
#### **4. DATA TRANSFORMATION**
<p>&nbsp;</p>

```{r } 
## in order to REMOVE the RECORDS where the GRADE 3 is > 2 :
## a new piece of R code (in contrast to the previous version)

dim(student3)

student3$G3 = as.integer(student3$G3)

student4 = student3[student3$G3 > 2, ]

dim(student4) 

ggplot(data = student4) + 
       geom_bar(mapping = aes(x=G3, fill=G3))

ggsave("display.0.G3.after.filtering.grade3.png")

student3 = student4

## TRANSFORMING G3 into RANGES of PASS and NO-PASS :

student3$G3 = as.integer(student3$G3)

student3$RESULT[student3$G3 <= 10] = "NO_PASS"
student3$RESULT[student3$G3 >=10 ] = "PASS"

student3 <- subset(student3, select = -c(G3))

student3$RESULT = as.factor(student3$RESULT)
```

<p>&nbsp;</p>
#### **5. TRAINING AND TEST SETS**
<p>&nbsp;</p>

```{r } 
## CHOOSING the TRAINING and TESTING SETS

indxTrain <- createDataPartition(student3$RESULT, 
                                 p = .75, 
                                 list = FALSE)

training <- student3[indxTrain,]
#training

testing <- student3[-indxTrain,]
#testing

dim(student3)
dim(training)
dim(testing)
```

<p>&nbsp;</p>
#### **6. PRE-PROCESSING THE DATA**
<p>&nbsp;</p>

```{r } 
## PRE-PROCESSING the DATA

trainX        <- training[, names(training) != "RESULT"]

preProcValues <- preProcess(x = trainX, method = c("center", "scale"))

preProcValues

names(trainX)

dim(trainX)

names(training)
```

<p>&nbsp;</p>
#### **7. PERFORMING THE TRAINING**
<p>&nbsp;</p>

```{r } 
## PERFORMING the TRAINING

set.seed(400)
ctrl <- trainControl(method="repeatedcv",repeats = 3) 

knnFit <- train( RESULT~ ., 
                 data = training, 
                 method = "knn", 
                 trControl = ctrl, 
                 preProcess = c("center","scale"), tuneLength = 20)

## The output of kNN fit

knnFit

png("the.results.knn.FIT.png")
plot(knnFit)
dev.off()

```

<p>&nbsp;</p>
#### **8. MAKING THE PREDICTIONS**
<p>&nbsp;</p>

```{r } 
## Making the PREDICTIONS :

knnPredict <- predict(knnFit, newdata = testing)

```

<p>&nbsp;</p>
#### **9. THE CONFUSION MATRIX**
<p>&nbsp;</p>

```{r } 
## COMPUTING the CONFUSION MATRIX :

confusionMatrix(knnPredict, testing$RESULT)

mean(knnPredict == testing$RESULT)

dim(student3)

```

<p>&nbsp;</p>

We may aim to optimize the model by feature selection or by including new features from the data that is available.

<p>&nbsp;</p>