---
title: "ASSOCIATION RULES"
author: "Bogdan Tanasa"
date: ""
output:
  pdf_document: default
  html_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<p>&nbsp;</p>

#### **1. DATA EXPLORATION**

<p>&nbsp;</p>

#### **2. ASSOCIATION RULES**

<p>&nbsp;</p>


\newpage
<p>&nbsp;</p>
#### **1. DATA EXPLORATION**
<p>&nbsp;</p>

&nbsp;
&nbsp;

We are using the data that we had from **UCI** a while ago in the file "Association_DataSet.csv".

```{r }
options(warn=-1)
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(arules))
suppressPackageStartupMessages(library(arulesViz))

######################################################
######################################################

FILE1="Association_DataSet.csv"

######################################################
######################################################

file = read.delim("Association_DataSet.csv", sep = ",", header=TRUE, stringsAsFactors=F)

######################################################
######################################################

str(file)
class(file)
summary(file) 

######################################################
######################################################

file$Family = as.factor(file$Family)
file$Hobbies  = as.factor(file$Hobbies)
file$Social_Club = as.factor(file$Social_Club)
file$Political  = as.factor(file$Political)
file$Professional = as.factor(file$Professional)
file$Religious  = as.factor(file$Religious)
file$Support_Group = as.factor(file$Support_Group)

summary(file) 

######################################################
######################################################
# to exclude Elapsed_Time and Age
# to transform 0 and 1 into YES or NO 

a_file <- file[, -which(names(file) %in% c("Elapsed_Time", "Age"))]

summary(a_file)

##  Time_in_Community     Gender            Working          Family   Hobbies 
##  Length:3483        Length:3483        Length:3483        0:2125   0:2438  
##  Class :character   Class :character   Class :character   1:1358   1:1045  
##  Mode  :character   Mode  :character   Mode  :character                    
##  Social_Club Political Professional Religious Support_Group
##  0:2828      0:3156    0:2353       0:2025    0:2930       
##  1: 655      1: 327    1:1130       1:1458    1: 553     

a_file$Family = ifelse(a_file$Family == "0", "Family_No", "Family_Yes")
a_file$Hobbies = ifelse(a_file$Hobbies == "0", "Hobbies_No", "Hobbies_Yes") 
a_file$Social_Club = ifelse(a_file$Social_Club == "0", "Social_Club_No", "Social_Club_Yes") 
a_file$Political = ifelse(a_file$Political == "0", "Political_No", "Political_Yes") 
a_file$Professional  = ifelse(a_file$Professional == "0", "Professional_No", "Professional_Yes") 
a_file$Religious = ifelse(a_file$Religious == "0", "Religious_No", "Religious_Yes") 
a_file$Support_Group = ifelse(a_file$Support_Group == "0", "Support_Group_No", "Support_Group_Yes") 

summary(a_file)

write.csv(a_file, file = "the_dataset.csv", row.names = FALSE)
```

&nbsp;
&nbsp;

\newpage
<p>&nbsp;</p>
#### **2. ASSOCIATION RULES**
<p>&nbsp;</p>
\newpage

Here we are performing the association analysis and we display the data.

```{r }
#####################################################################################

the_data <- read.transactions("the_dataset.csv", sep = ",", header=TRUE)

##################################################################
##################################################################

summary(the_data)
# inspect(the_data)
inspect(the_data[1:5])

##################################################################
##################################################################

itemFrequencyPlot(the_data, support = 0.1)

itemFrequencyPlot(the_data, topN = 20)

##################################################################
##################################################################

# to visualize the data :

image(sample(the_data, 100))

##################################################################
##################################################################

# if we attempt to use the default settings of support = 0.1 and confidence = 0.8, 
# find a set of 2918 rules:

the_rules = apriori(the_data)

summary(the_rules)

# if we change the settings we may find less rules:

# the_rules = apriori(the_data, parameter = list(support = 0.1, 
#                                               confidence = 0.8, 
#                                               minlen = 2))

# summary(the_rules)
```

\newpage
<p>&nbsp;</p>
**STRONGLY SUPPORTED ASSOCIATION RULES**
&nbsp;
\newpage

```{r }
################################################################## DISPLAYING strongly 
################################################################## SUPPORTED RULES

rules.sorted <- sort(the_rules, by="lift")

# inspect(rules.sorted)

inspect(rules.sorted[1:5])

## saving the data :

write(the_rules, file = "the_rules.csv", 
                 sep = ",", 
                 quote = TRUE, 
                 row.names = FALSE)
```

\newpage
<p>&nbsp;</p>

```{r }
## visualizing the data :

plot(the_rules)

plot(the_rules, method="graph", control=list(type="items"))

# plot(the_rules, method="paracoord", control=list(reorder=TRUE))

############################################################################
############################################################################
################################ a piece of R code if we may have to use
################################ depending on the context

# inspect(rules.sorted)

# prune redundant rules.
# subset.matrix <- is.subset(rules.sorted, rules.sorted)
# subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
# redundant <- colSums(subset.matrix, na.rm=T) >= 1
# which(redundant)

# remove redundant rules.
# rules.pruned <- rules.sorted[!redundant]
# inspect(rules.pruned)
```

\newpage